---
title: "Descriptive Analysis of CSSS Data"
date: 24th July, 2019
author: Fabian Dablander
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = 'center')
```


# Data Preparation
```{r}
library('dplyr')
library('papaja')
library('ggplot2')
library('SentimentAnalysis')


get_sentiment <- function(text) {
  analyzeSentiment(text)$SentimentQDAP
}


# datd <- read.csv('data/csss-dakota.csv', na.strings = '')
# datf <- read.csv('data/csss-fabian.csv', na.strings = '')
# datj <- read.csv('data/csss-jackie.csv', na.strings = '')
# date <- read.csv('data/csss-emily.csv', na.strings = '')
# datk <- read.csv('data/csss-kyle.csv', na.strings = c('NA', 'NA1', 'NA2'))

# add sentiment of biography because ... why not?
# dat <- rbind(datf, date, datd, datk, datj) %>% 
#   mutate(
#     Bio_sentiment = get_sentiment(as.character(Biography)),
#     Iteration = interaction(Year, Location),
#     Country_University = as.character(Country_University),
#     Country_University = ifelse(
#       Country_University %in% c('US', 'USA', 'United States'), 'USA', Country_University
#     )
#   )

dat <- read.csv('data/cleaned_csss-all.csv', na.strings = '') %>% 
  mutate(
    Bio_sentiment = get_sentiment(as.character(Biography)),
    Iteration = interaction(Year, Location),
    Country_University = as.character(Country_University),
    Country_University = ifelse(
      Country_University %in% c('US', 'USA', 'United States'), 'USA', Country_University
    )
  )
```


# Data Analysis
## Descriptive Statistics
Here we look at a few summary statistics and how they have changed over the years.

```{r Summary-Statistics, results = 'asis'}
dat_avg <- dat %>%
  group_by(Year, Location) %>% 
  summarize(
    prop_female = mean(Gender == 'Female', na.rm = TRUE),
    avg_group_size = mean(Size, na.rm = TRUE),
    prop_US = mean(Country_University %in% c('US', 'USA', 'United States')),
    nr_projects = length(unique(Title)),
    nr_participants = length(unique(Name)),
    nr_projects_per_part = nr_participants / nr_projects,
    prop_PhD = mean(Position == 'Student', na.rm = TRUE), # this overcounts; fix!
    prop_postdoc = mean(Position == 'PostDoc', na.rm = TRUE),
    prop_industry = mean(Position == 'Industry', na.rm = TRUE),
    avg_sentiment = mean(Bio_sentiment, na.rm = TRUE)
  ) %>% 
  mutate(
    Iteration = factor(interaction(Year, Location))
  )

tab_names <- c(
  'Year', 'Location', '%Female', 'Avg_Group_Size', '%US', '#Projects',
  '#Participants', '#Proj_per_Part', '%PhD', '%PostDoc', '%Industry', 'Avg_Sentiment'
)
apa_table(setNames(dat_avg, tab_names))
```

Note that the summer school of 2011 apparently had only one project in which everybody took part. The percent of women participants has become close to what it should be in the last two years. There is some issues with the years 2012-2014 with respect to the number of PhD students.

## Proportion of Females across Time
```{r}
ggplot(filter(dat_avg, Iteration != '2005.Beijing'), aes(x = Iteration, y = prop_female)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  theme_bw() +
  ylab('% Women') +
  xlab('') +
  scale_y_continuous(limits = c(0, .6), breaks = scales::pretty_breaks(n = 10)) +
  geom_hline(yintercept = .5, linetype = 'dashed') +
  ggtitle('Proportion of Women across CSSS Iterations') +
  theme(
    axis.text.x = element_text(angle = 90, size = 10),
    axis.text.y = element_text(size = 14),
    axis.title.x = element_text(size = 14),
    plot.title = element_text(hjust = .5, size = 16)
  )
```

## Disciplines across Time
We need to be careful since some people can come from more than one discipline.
```{r}
library('stringr')

create_disciplines_dat <- function(dat) {
  dat_disciplines <- c()
  iterations <- unique(dat$Iteration)
  
  for (school in iterations) {
    disciplines <- str_split(filter(dat, Iteration == school)$Discipline_isced, ';|,', simplify = TRUE)
    disciplines <- str_to_title(str_trim(disciplines))
    disciplines <- disciplines[disciplines != '']
    
    if (!all(is.na(disciplines))) {
      tab <- table(disciplines)
      d <- cbind(school, data.frame(tab), as.numeric(tab) / sum(as.numeric(tab)))
      dat_disciplines <- rbind(dat_disciplines, d)
    }
  }
  
  dat_disciplines
}

dat_disciplines <- create_disciplines_dat(dat)
```

```{r, fig.width = 10, fig.height = 14}
colnames(dat_disciplines) <- c('Iteration', 'Disciplines', 'Frequency', 'Proportion')
uniq_disciplines <- as.character(unique(dat_disciplines$Disciplines))
sorted_levels <- uniq_disciplines[order(nchar(uniq_disciplines), uniq_disciplines, decreasing = TRUE)]
dat_disciplines$Disciplines <- factor(dat_disciplines$Disciplines, levels = sorted_levels)

ggplot(dat_disciplines, aes(x = Disciplines, y = Proportion)) +
  geom_bar(stat = 'identity') +
  facet_wrap(~ Iteration) +
  coord_flip() +
  theme_bw() +
  ylab('') +
  theme(axis.text.x = element_text(angle = 90, size = 6))
```

```{r}
iter <- c('2006.Santa Fe', '2010.Santa Fe', '2018.Santa Fe')
dat_disciplines_selection <- filter(dat_disciplines, Iteration %in% iter)
dat_disciplines_selection$Iteration <- factor(dat_disciplines_selection$Iteration, levels = iter)

ggplot(dat_disciplines_selection, aes(x = Disciplines, y = Proportion)) +
  geom_bar(stat = 'identity') +
  facet_wrap(~ Iteration) +
  coord_flip() +
  theme_bw() +
  ylab('Proportion') +
  xlab('') +
  ggtitle('Prevalence of Disciplines across three Years') +
  theme(
    axis.text.x = element_text(angle = 90, size = 8),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(hjust = .5, size = 16)
  )
```

## Country across Time
```{r}
euro.countries <- c(
  "Austria", "Belgium", "Czech Republic", "Denmark", "Estonia", "France",
  "Germany", "Hungary", "Iceland", "Ireland", "Italy", "Netherlands", "Norway",
  "Poland", "Portugal", "Russia", "Serbia", "Serbia and Montenegro", "Slovenia",
  "Spain", "Sweden", "Switzerland", "The Netherlands", "Turkey", "UK", "United Kingdom"
)

afri.countries <- c("Nigeria", "South Africa")
asian.countries <- c(
  "China", "India", "Iran", "Israel", "Japan", "Lebanon", "Hong Kong",
  "Pakistan", "South Korea", "Singapore", "Signapore", "Philippines"
)

s.amer.countries <- c("Argentina", "Argentinia", "Brazil", "Colombia", "Columbia", "Chile", "Venezuela")
n.amer.countries <- c("Canada", "Canda", "Costa Rica", "Guatemala", "Mexico", "United States", "US", "USA")

get_continent <- function(country) {
  if (country %in% euro.countries) {
    res <- 'Europe'
  } else if (country %in% afri.countries) {
    res <- 'Africa'
  } else if (country %in% asian.countries) {
    res <- 'Asia'
  } else if (country %in% s.amer.countries) {
    res <- 'South America'
  } else if (country %in% n.amer.countries) {
    res <- 'North America'
  } else if (country %in% 'Australia') {
    res <- 'Australia'
  } else if (country %in% 'New Zealand') {
    res <- 'New Zealand'
  } else {
    res <- NA
  }
  
  res
}

dat <- dat %>% 
  mutate(
    Continent_University = Vectorize(get_continent)(Country_University),
    Continent_University = factor(
      Continent_University, levels = sort(unique(Continent_University), decreasing = TRUE)
    )
  )

ggplot(dat, aes(x = Continent_University)) +
  geom_bar() +
  facet_wrap(~ Iteration) +
  coord_flip() +
  theme_bw() +
  ylab('') +
  theme(axis.text.x = element_text(angle = 90, size = 6))
```

## Gini Coefficient
```{r}
library('DescTools')

DISCIPLINES <- as.character(unique(dat_disciplines$Disciplines))
map <- list()

i <- 1
for (d in DISCIPLINES) {
  map[[d]] <- i
  i <- i + 1
}

dat_dsum<- dat_disciplines %>% 
  mutate(
    Disciplines_int = map[as.character(Disciplines)]
  ) %>% 
  group_by(Iteration) %>% 
  summarize(gini = Gini(Disciplines_int)) %>% 
  mutate(
    Iteration = factor(Iteration, levels = sort(as.character(Iteration)))
  )

ggplot(dat_dsum, aes(x = Iteration, y = gini)) +
  geom_point(stat = 'identity') +
  geom_line(group = 1) +
  theme_bw() +
  xlab('') +
  ggtitle('Inequality of Discipline Prevalence across Iterations') +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10), limits = c(0, 1)) +
  ylab('Gini Coefficient') +
  theme(
    axis.text.x = element_text(angle = 90, size = 12),
    axis.title.y = element_text(angle = 90, size = 14),
    plot.title = element_text(hjust = .5, size = 16)
  )

ggplot(dat_dsum, aes(x = Iteration, y = gini)) +
  geom_bar(stat = 'identity') +
  theme_bw() +
  xlab('') +
  coord_flip() +
  ggtitle('Inequality of Discipline Prevalence across Iterations') +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10), limits = c(0, .6)) +
  ylab('Gini Coefficient') +
  theme(
    axis.text.x = element_text(angle = 90, size = 10),
    axis.text.y = element_text(size = 14),
    axis.title.x = element_text(size = 16),
    plot.title = element_text(hjust = .5, size = 16)
  )
```

## Survey Responses
First, we rename the columns to have more concise names.

```{r}
new_names <- c(
  'Timestamp', 'Name', 'Year_Incorrect', 'Year', 'Location', 'Discipline', 'Discipline_isced',
  'Closeness_Projects', 'Communication_Difficulty', 'Input_Usefulness', 'CSSS_Influence', 'Published',
  'Publishing_Difficulty', 'Continued_Collaboration', 'Interdisciplinary_Thoughts'
)

choose_first <- function(disciplines) {
  short_disc <- str_split(disciplines, ';|,', simplify = TRUE)
  short_disc[, 1]
}

dats <- read.csv('data/csss-survey.csv')
colnames(dats) <- new_names
dats <- dats %>% 
  mutate(
    Published = ifelse(Published == 'Yes', 1, 0),
    Continued_Collaboration = ifelse(Continued_Collaboration == 'Yes', 1, 0),
    Iteration = interaction(Year, Location),
    First_Discipline = factor(choose_first(Discipline_isced), 
  )

dats_disciplines <- create_disciplines_dat(dats)
```

## CSSS Influence on Discipline
The figure below shows how strongly different disciplines have been influenced by the CSSS summer school. Note that I only take the first discipline of each person as an approximation. Regardless, it seems that the social and behavioural sciences, as well as the life sciences, are more influenced then the physical sciences.

```{r}
ggplot(filter(dats, First_Discipline != ''),
       aes(x = First_Discipline, y = CSSS_Influence, fill = First_Discipline)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitter(width = .3, height = .05), size = .6, color = 'black') +
  theme_bw() +
  ylab('Influence of CSSS') +
  xlab('') +
  ggtitle('Influence of the CSSS by Discipline') +
  theme(
    axis.text.x = element_text(angle = 90, size = 12),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(size = 14),
    plot.title = element_text(hjust = .5, size = 16),
    legend.position = 'None'
  ) +
  coord_flip()
```

We can make the same plots with the usefulness of input, continued collaboration, and communication difficulty.
```{r}
ggplot(filter(dats, First_Discipline != ''),
       aes(x = First_Discipline, y = Input_Usefulness, fill = First_Discipline)) +
  geom_boxplot() +
  geom_point(position = 'jitter') +
  theme_bw() +
  ylab('Usefulness of Input') +
  xlab('Discipline') +
  theme(
    axis.text.x = element_text(angle = 90, size = 12),
    legend.position = 'None'
  ) +
  coord_flip()

dat_col <- filter(dats, First_Discipline != '') %>%
  group_by(First_Discipline) %>%
  summarize(
    mean_collaboration = mean(Continued_Collaboration),
    sd_collaboration = sd(Continued_Collaboration)
  )
    

# Violin plots to see the bimodality better
ggplot(dat_col, aes(x = First_Discipline, y = mean_collaboration, color = First_Discipline)) +
  geom_point() +
  geom_errorbar(
    aes(
      ymin = ifelse(mean_collaboration - sd_collaboration < 0, 0, mean_collaboration - sd_collaboration),
      ymax = ifelse(mean_collaboration + sd_collaboration > 1, 1, mean_collaboration + sd_collaboration)
    )
  ) +
  theme_bw() +
  ylab('% Continued Collaboration') +
  xlab('Discipline') +
  theme(
    axis.text.x = element_text(angle = 90, size = 12),
    legend.position = 'None'
  ) +
  coord_flip()


ggplot(filter(dats, First_Discipline != ''),
       aes(x = First_Discipline, y = Communication_Difficulty, fill = First_Discipline)) +
  geom_boxplot() +
  geom_point(position = 'jitter') +
  theme_bw() +
  ylab('Difficulty of Communication') +
  xlab('Discipline') +
  theme(
    axis.text.x = element_text(angle = 90, size = 12),
    legend.position = 'None'
  ) +
  coord_flip()

ggplot(filter(dats, First_Discipline != ''),
       aes(x = First_Discipline, y = Communication_Difficulty, fill = First_Discipline)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitter(width = .3, height = .05), size = .6, color = 'black') +
  theme_bw() +
  ylab('Difficulty of Communication') +
  xlab('') +
  ggtitle('Difficulty of Communication by Discipline') +
  theme(
    axis.text.x = element_text(angle = 90, size = 12),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(size = 14),
    plot.title = element_text(hjust = .5, size = 16),
    legend.position = 'None'
  ) +
  coord_flip()
```

### Word cloud
```{r}
library('tm')
library('wordcloud')
library('RColorBrewer')

docs <- Corpus(VectorSource(as.character(dats$Interdisciplinary_Thoughts)))
docs <- tm_map(docs, removePunctuation)
docs <- tm_map(docs, removeWords, stopwords('english'))

dtm <- TermDocumentMatrix(docs)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)

wordcloud(words = d$word, freq = d$freq, min.freq = 1,
          max.words = 100, random.order = FALSE, rot.per = 0.35, 
          colors = brewer.pal(8, "Dark2"))
```


```{r}
"5 stars, would recommend"
"Useful from an academic perspective, less useful from a professional perspective"
"Difficult to get credited for it by traditional systems of evaluation."
"Essential. Not always straight-forward, but totally worth the effort"
"Best thing in the world: I still have life-long friendship and collaboration with SFI pals!"
"Very necessary if the right attitude is present. Multidisciplinary is not just about disciplines coming together 'because'. The scientific question should come first."
"Hihgly desirable but the academic system favours specialization over insterdisciplinary research"
"You asked about group projects. I don't recall group projects back in 1990."
"Necessary for me to be happy"
```


